\documentclass[answers,12pt,addpoints]{exam} 
\usepackage{import}
\usepackage{hyperref}

\import{C:/Users/prana/OneDrive/Desktop/MathNotes}{style.tex}

% Header
\newcommand{\name}{Pranav Tikkawar}
\newcommand{\course}{StockBee}
\newcommand{\assignment}{Day 1}
\author{\name}
\title{\course \ - \assignment}

\begin{document}
\maketitle
\tableofcontents
\newpage

\section{Session 1}

\textbf{Formula based:} without it, it's pure random.\\
\textbf{2 ways to make money:}
\begin{itemize}
    \item Focus on Larger Moves
    \item Focus on Smaller Moves
\end{itemize}
Most trader focus on smaller moves: \textbf{Scalping}, \textbf{Day Trading}, \textbf{Swing Trading}.\\
With a bankroll of 100k you look for 500 to 5000 dollar moves: HFT.\\
One dude does 6M trades in day, with average .1 cent profit per trade, making 6k a day.\\
\textbf{Today: Larger Moves}\\
Characterized by a large magnitude moves.\\\\
\textbf{NEWS BRING Moves:} Not every thing is contained in the chart.\\
\textit{Question: What news sources, and how do you find them?}\\
Become an expert on News and Catalysts\\
The market has "nothing" to do with the economy, it moves based on sentiment\\
We are traders not investors. Capitalize on the hope and fear of the market.\\
Avoid overovervalued stocks as one issue means hell.\\
\textit{Question: How to verify the news? How to know if it's real?}: Check earnings, SEC filings, etc. Look at the briefing briefing.com/ the fly https://thefly.com/news.php \\
\textbf{News Catalysts: 2 types}
\begin{itemize}
    \item Real/Fundamental Catalysts: Characterized by actual sales, earnings, and news. Look at companies that announce earnings.
    \item Story Catalysts: Characterized by rumors, hype, and speculation. Look at companies that are hyped up.
\end{itemize}
For every 1 real news, there are 300+ story catalysts.\\
Learn to look at the X-Ray of the company.\\
\textbf{Larger Moves:} 10\%+ moves.\\
Largest moves is the sales growth. IE New product\\
Profit Growth: not as good but still good: Good for a turn around situation\\
Government Policy changes: IE Trump and China\\
Big orders: \\
Management Changes: CFO Leaving = Bad\\
Diseases and Disaster:\\
Think what is the trade in this.\\
Best way to test is to do this by looking at news \\
Need to convert to a daily watchlist. News based 7/11.\\
Into a Checklist \textbf{MAGNA53 Cap10*10}\\
ntrt mtrt: night time is right time, morning time is right time\\
\begin{itemize}
    \item MA: massive sales growth/earnings growth. Massive acceleration in projections
    \item G: Gap up. in pre/post market. +4\% plus 100k volume 
    \item N: neglect. (impt!!!) price, volume, news, funds, analysts 
    \item A: acceleration in sales
    \item 5: 5 days plus short interest it will takes 5 days to cover
    \item 3: 3 plus analyst upgrades? (\textit{Explain later}) 
    \item Cap10*10: Get rid of stocks with market cap over 10B. 
\end{itemize}
IPO in past 10, low market cap ($<$1B), 
\textit{Question: How do I know where some of these scans come from. How do I test my own?}\\
\section{Session 2}
Formula based trading Over emotions and randoms.\\
Structural stop: Moving stoploss based on the structure of the stock.\\
Episodic Pivot: Why? trajectory change. It throws the audience for a loop.\\
\textbf{Episodic Pivot:} \textit{Ask Later} \\
Two major types of Magna53 stocks: 
\begin{itemize}
    \item Growth Stocks: New industry, new technology, consumer discretionary: healthcare, biotech, 
    \item Turnaround stocks: Old industry, old technology, consumer staples.
\end{itemize}
Q over Q sales growth, sustainable, mean its a good call to buy.\\
Bucket: \textit{Kinda lost for the buckets}
\begin{itemize}
    \item 0 39\% 2q in a row. cap 10*10. base minimum sales is 20M
    \item cap <10B ipo <10 years 39\% 2q in a row
\end{itemize}
Turnaround stocks can make a larger move,
\begin{itemize}
    \item Cyclical industry
    \item Consumer discretionary
\end{itemize}
when company lost alot of interest and then makes a comback. Bearish market then going up.\\
They tend to last for 2-3 years\\
We might see 5-6 growth stocks in a year.\\
We might see 20-30 turnaround stocks in a year.\\
EP 9 million. Delayed EP 9 million.\textit{Explain Later}\\
Sudden volume surge of 10 mil plus is a catalyst\\
\section{Session 3}
EP vol $>$8.9
\textit{Review this later}
\begin{itemize}
    \item Cat (catalyst) 
    \begin{itemize}
        \item Real estate
        \item size
    \end{itemize}
    \item Dog
    \begin{itemize}
        \item story
        \item 100 million 
    \end{itemize}
    \item Liquid Lava
    \begin{itemize}
        \item liquid
        \item size
    \end{itemize}
\end{itemize}
slow moves persist, fast moves fade.\\
There are alot of opportunities in EP 9M. So we can use selection methods on top of that.\\
Priotitization:
\begin{enumerate}
    \item lowest cap
    \item highest volume
    \item youngest move
\end{enumerate}
300 - 500 opportunities\\
Delayed EP 9M: $< 5\%$ stop loss\\
Characterists of stocs that EP 9M: have swings of 20-40 \% \\
Sugar babies: they give you alot more fun!\\
High risk high reward.\\

\section{Day 1 Summary}
money = formula: replicable and simple\\
2 kinds of formula: bigger moves, smaller moves\\
Bigger Moves 20\% - 100\%\\
Catalysts/news: news makes moves\\
smaller moves 1\% - 5\%\\
ntrt mtrt\\
EP: Episodic Pivot \\
Growth opportunities: growth stocks, tech, consumer discretionary\\
Turnaround opportunities: cyclical industry, consumer discretionary\\
EP 9M: Volume is a catalyst\\
Convert to cats dogs and liquid lava\\
trade cats and liquid lava with large size becasue of conviction\\
trade dogs with small size because they r dogdy \\
Lowest cap, highest volume, youngest move\\
Delayed EP 9M: is an anticipated move. Consolidation and pullback.\\
\section{Day 2: Session 4}
First thing to do after scan to do X-Ray \\
Find Reason for trades and sizing \\
More you understand the field, the better you can trade.\\
Today is \textbf{Pattern Based Trading}.
Focus on Procedural Trading Patterns. No need to even think about random events.\\
Buying implies more buying, and selling implies more selling.\\
Momentum burst is a well studied phenomenon.Stock goes up 3-4 days very likely to have 1 day pullback.\\
\textit{Question: How do I formulate the notion of momentum, and changes of 3-5 days in a formula}\\
Stocks move in bursts of 3-5 days, 8-40\% magnitutde, They require no specific catalyst. Usually happen near Yearly high low, and all time high low. \\
It is an inherent tendency of the market.\\
SOS - start of swing. It is a signal \\
SOS continuation\\
SOS consolidation - Catalysts and volume. Best will be EP, EP9M\\
Favorite catalyst is sales growth\\
SOS bottom bounce - T2108 when below 20\\
2lynch
\begin{itemize}
    \item 2 - Not up 2 days in a row
    \item L - Linear, first leg is consistent and persistent buying ie strong demand
    \item Y - young, 1st, 2nd leg move
    \item N - Narrow/Negative day prior to breakout
    \item C - Consolidation quality. Orderly 4\% breakdown. low volume, low volatility prior
    \item H - Closing near highd 
\end{itemize}
Consolidation:
\begin{itemize}
    \item Start of swing 
    \item Catalyst
    \item Volume
\end{itemize}
No need to trade choppy moves. And then we need wide stops: drama. \\
High price, low volatility, but want to be in the middle to get good returns and minimize risk. \\
If it meets 2lynch but not H its an anticipation setup \\
For singles:
\begin{itemize}
    \item SOS
    \item Continuation
    \item Consolidation
    \item Bottom Bounce
\end{itemize}
Inherent market tendencies
\begin{itemize}
    \item Momentum burst
    \item Exhaustion
\end{itemize}
Focus on Position sizing, IF know 1 day goes up, 2 day goes, up 3 is volatitle. Large position day 1, remove 80\% on day 2. stop loss for day 3.\\
Anticipation:
\begin{itemize}
    \item SOS
    \item WSS: weak structure stock
    \item Parabolic Short (risky)
\end{itemize}
WSS: Niagara 3*10, weak 3*2.5\\
look for stock down by 10\% in 3 days\\
cannot bounce by 2.5\% at the consolidation\\
Most downtrend is 3 days 
Process Orientation
Focus on the checklist and timing 
Pressure builds diamonds. \\

\section{Session 5}
Arthur Bandura
Self-efficacy belief and mastery experience.\\
Vicarious experience.\\
Social persuasion.\\
Visualization.\\
psycho-somatic state
Emotional and physiological states.\\
All of this is do able by yourself and changing your beliefs\\
Overtrading is a big issue.\\

\newpage
\section{Review}
\textbf{1: Momentum}\\
\textbf{2: Short term} $\rightarrow 3 -5$ days\\
\textbf{3: Setup}
\begin{itemize}
    \item Continuation
    \item Consolidation
    \item Bottom Bounce
    \item EPs
    \begin{itemize}
        \item 9M 
        \item Catalyst
    \end{itemize}
    \item SIPS: stocks in play
\end{itemize}
Continuation: \\
1st scan\\
2nd 2lynch: quality criteria\\
when stocks move they move in bursts 3-5 days.\\
3rd Execution:ie risk management and stop loss\\
Anticipation: \\
SOS: risky\\
Consolidation setup:\\
Low volatility (range bound)  \\
Then a catalyst (news) to make it go up \\
Episodic Pivot:\\
9M EP: 9M volume
\textit{Look into significant volume spike}
Other EPS: but later issue (earnings based): ntrt mtrt \\
SIPS: stocks in play: IE short term new bursts
\textit{Anticipate: News Sites for a scraper for news}

\textit{Next time: how is scan write}

Factor investing\\
seeminly  uncorrleated thing to predict stock\\
Simple factors: Inser tradsing\\ 
sum os indersing selling at the same time\\
This is an Alpha: with easily available data \& it needs to be compiled properly
\textbf{scraper} and then \textbf{look for patterns}\\
Some time when the insders sell then stock goes down\\
Other PEADs post earning announcement drift.\\
When stocks come out with earnings and earnings are significant then stock drifts in terms of earnings\\
Its difficult to look at all. \\
\textbf{build scaners, for earning which beats expectings, sales are good, priortize those } Those make amazing moves.\\
Start with basic models: \\
Sophisticated: Seasonaility: show all stock with DATE sznality. \\
IE how to remove the seasonality with persistenc and then rank them by how valuable and consistent. \\
Pairs trading: Goog goes up and other stocks also move. \\
Cluster of factors that explain why XYZ. \\
Whiteness market vs Stain remover makrket. Focus on audiance and "why" the audience wants XYZ.\\
Using the cluster and correlation for factors for the scans:
\begin{itemize}
    \item Factor based trading: \emph{info is public but difficult to use}
    \begin{itemize}
        \item Insider Trading: Bar Chart and Kalashi.
        \begin{itemize}
            \item Look into the valuable people and see if it correlates to the stock price. IE CFO sell $\implies$ stock down
            \item IDEA: ppl on insde have more info
        \end{itemize}
        \item PEADs: ***
        \begin{itemize}
            \item Look into financials.
            \item Look at where the analysts expecting: \textit{Where do they post these}
            \item Do we beat public sentiment and if no then it will likely to move towards that.
        \end{itemize}
    \end{itemize}
\end{itemize}
Start with the first 2. \\
Make it modular and visualizaible \\
Emph for the user to actually use. \\
Have buckets and order the stocks on how many buckets they fill. \\
UI: show the top bucets the top. Show the Anomaly\\ 
\textit{Look for 10K, 10Q, other fin statment schedule}. There exists ad earnings calendar\\
The statments are usually sent before-or after makret. \\
We are interested in the extreme: Good or Ugly not the usually. Anomaly. \\
EDGAR API\\
Not just to have data, what does it mean.\\
Earning Whisper, Zacks. \\
Objective to make money.\\
Make a whiteborad.\\
\textbf{look what they do in Quant.}






\end{document}