\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{mathrsfs}
\usepackage{cancel}

\usepackage{graphicx}


\setlength\parindent{0pt}

\author{Pranav Tikkawar}
\title{Assignment 1}

\begin{document}
\maketitle
\section*{Question 7}
\subsection*{a}
$\frac{6!}{3!3!}$ as there are 6 people in total to arrange with 3 and 3 indistinguishable boys and girls.\\
But if the boys and girls are distguishable (meaning they are actually different human) then the outcome is $6!$.
\subsection*{b}
$2!$ as we can consider the boys and girls blocks, leaving 2 options(in the case that they are indistinguishable).\\
If they are distinguishable then the outcome is $2! \cdot 3! \cdot 3!$ as there are 2 blocks of boys and girls that can be arranged in $3!$ ways.  
\subsection*{c}
$4!$ as we can consider the boys are a block (and indistinguishable) and thus we are left $4$ items to arrange.\\
If the boys are distinguishable then the outcome is $4! \cdot 3!$ as there are 4! ways to arrange the block and then 3! ways to arrange the boys inside the boys block.
\subsection*{d}
$3! \cdot 3! \cdot 2$ as there are 3! ways to order the boys and girls and 2 ways to start of with either a boy or girl. The outcomes is the same if they children are distinguishable or indistinguishable.
\section*{Question 10}
\subsection*{a}
$8!$ as there are 8 spots with no restrictions.
\subsection*{b}
$7! \cdot 2$ as the block of A and B is treated as one spot leading to 7 spots and then there are 2 ways to organize the block of 2 people.
\subsection*{c}
$4! \cdot 4! \cdot 2$ as there are 4! ways to order the men and the women, and there are 2 different ways to start.
\subsection*{d}
$5! \cdot 4!$ as the 5 men become 1 block leaving 4 items, and then there are $4!$ ways to order those for blocks and then there are $5!$ ways to order the men in the block of 5 men.
\subsection*{e}
$4! \cdot 2^{4}$ as there are 4! ways to order the couples and then 2! ways to order each of the couples leading to $2^{4}$ ways to order all the couples.
\section*{Question 11}
\subsection*{a}
$\frac{6!}{3!2!}$ as there are 6 books and then divide by the number of ways to order the 3 and 2 indistinguishable books. \\
If the books are distinguishable then the outcome is $6!$
\subsection*{b}
$3!$ as the books become 3 block and since each item in the block is indistinguishable, there are 3! ways to order the blocks.\\
If the books are distguishable then the outcome is $3! \cdot 2! \cdot 3!$ as there are $3!$ ways to order the blocks and then $2!$ ways to order the books in the block of 2 and $3!$ ways to order the books in the block of 3.
\subsection*{c}
$4!\cdot 3!$ since the 3 novels become a block with there being 4 blocks in total which can be ordered $4!$ different ways and then $3!$ ways to order the books inside the block of 4.
\section*{Question 17}
$\frac{10!}{3!}$ as there are 10 spots for children with 7 that are distguishable with distinct gifts and 3 that are indistinguishable with no gifts. Thus it is $\frac{10!}{3!}$.


\end{document}